\modLP{Identification of} Biological Regulatory Networks from Process Hitting models

### 2014-10-25 parts-phage-lambda.tex

\modMF{
\subsection{Detailed example on the bacteriophage lambda immunity control model}
}
\label{ssec:phage-lambda}

@@

  % Sorte coopérative
  \newcommand{\ttable}[3]{\scriptsize%
    $#1\left\{\begin{array}{@{}l}\cI #2 1\\\cro #3 2\end{array}\right.$}
  \TSetTick{cIcro}{0}{\ttable{FF}{<}{<}}
  \TSetTick{cIcro}{1}{\ttable{FT}{<}{\leq}}
  \TSetTick{cIcro}{2}{\ttable{TF}{\leq}{<}}
  \TSetTick{cIcro}{3}{\ttable{TT}{\leq}{\leq}}
  \TSetSortLabel{cIcro}{$\cIcro$}
  \TSort{(4,0.5)}{cIcro}{4}{r}

@@

We first note that in the case where $\cro_3$ is present, we have from \pref{eq:ctx-sigma}:
\begin{align*}
  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \cI_0, \cro_3, \cIcro_{FT} \} \\
  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \cI_1, \cro_3, \cIcro_{TT} \} \\
  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \cI_2, \cro_3, \cIcro_{TT} \}
\end{align*}
We can then deduce the following directions of evolution:
\begin{align*}
  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} &&
    \text{as } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) &&
    \text{(from \pref{eq:B-no-bounce})} \\
  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} && &&
    \text{(from \pref{eq:B-bounce})} \\
  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \} && &&
    \text{(from \pref{eq:B-bounce})}
\end{align*}
Therefore, \pref{eq:inference-edges-ba} gives: $\cI \xrightarrow{1} \N \in \hat{E}_-$.
The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
do not add more information on the influences towards $\N$.
Thus, given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cI}{-}{1}{\N} \in E$.

@@

Let us first focus on the group of regulators $\{ \cro \}$.
Of course, $\forall \cro_i \in \PHl_{\cro}, \ctx_{\{ \cro \}}(\PHetat{\cro_i}) = \{ \cro_i \}$.
However, we note that all processes of $\cI$ can hit $\cro_1$ and $\cro_2$;
therefore $\Phi(\cro_1) = \Phi(\cro_2) = \emptyset$, and:
%, there exist no sub-state $\sigma \in \Phi(\cro_1)$
\begin{align*}
  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_3}) &= \{ \cro_2 \} && &&
    \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_2}) &= \emptyset &&
    \text{because } \Phi(\cro_2) = \emptyset &&
    \text{(from \pref{eq:B-no-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_1}) &= \emptyset &&
    \text{because } \Phi(\cro_1) = \emptyset &&
    \text{(from \pref{eq:B-no-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_0}) &= \{ \cro_0 \} &&
    \text{because } \PHetat{\cI_2, \cro_0} \in \Phi(\cro_0) &&
    \text{(from \pref{eq:B-no-bounce})}
\end{align*}

We thus cannot infer any local influence by observing the group of regulators $\{ \cro \}$.
Let us now focus instead on the group of regulators $\{ \cI \}$.
One again, we trivially have:
$\forall \cI_i \in \PHl_{\cI}, \forall \cro_j \in \PHl_{\cro},
  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_i, \cro_j}) = \{ \cI_i, \cro_j \}$.
Furthermore:
\begin{align*}
%   \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_0}) &= \{ \cro_1 \} &&
%   \text{(from \pref{eq:B-bounce})} \\
%   \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_1}) &= \{ \cro_2 \} &&
%   \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_2}) &= \{ \cro_3 \} &&
    \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3}) &= \{ \cro_2 \} &&
    \text{(from \pref{eq:B-bounce})}
\end{align*}
Therefore, \pref{eq:inference-edges-aa} gives: $\cro \xrightarrow{3} \cro \in \hat{E}_-$.
% The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
% do not add more information on the influences towards $\N$.
Finally, as no more information is added by observing the other levels of $\cro$ and $\cI$,
and given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cro}{-}{3}{\cro} \in E$.

@@

\begin{table}[t]
\begin{align*}
  K_{\cI, \emptyset} &= \{2\} &
  K_{\cI, \{\cro\}} &= \{0\} \\
  K_{\cI, \{\cII\}} &= \{2\} &
  K_{\cI, \{\cII;cro\}} &= \{2\} \\
  K_{\N, \emptyset} &= \{1\} &
  K_{\N, \{\cro\}} &= \{0\} \\
  K_{\N, \{\cI\}} &= \{0\} &
  K_{\N, \{\cI;cro\}} &= \{0\} \\
  K_{\cII, \emptyset} &= \{0\} &
  K_{\cII, \{\cro\}} &= \{0\} \\
  K_{\cII, \{\cI\}} &= \{0\} &
  K_{\cII, \{\cI;cro\}} &= \{0\} \\
  K_{\cII, \{\N\}} &= \{1\} &
  K_{\cII, \{\N;cro\}} &= \{0\} \\
  K_{\cII, \{\N;cI\}} &= \{0\} &
  K_{\cII, \{\N;cI;cro\}} &= \{0\} \\
  K_{\cro, \emptyset} &= \{3\} &
  K_{\cro, \{\cro\}} &= \{2\} \\
  K_{\cro, \{\cI\}} &= \{0\} &
  K_{\cro, \{\cI;cro\}} &= \{2\}
\end{align*}
\caption{\label{tb:phage-lambda-k}%
  Result of the parameters inference performed on the Process Hitting model
  of the phage lambda immunity response
  taken from \cite{thieffry_dynamical_1995}.
  \todo{Remove the curly brackets $\{\}$?}
}
\end{table}

>>>>

\modMF{
\subsection{Detailed example: the bacteriophage lambda immunity control}
}
\label{ssec:phage-lambda}

@@

  % Sorte coopérative
  \newcommand{\ttable}[3]{\scriptsize%
    $#1\left\{\begin{array}{@{}l}\cI #2 1\\\cro #3 2\end{array}\right.$}
  \TSetTick{cIcro}{0}{\ttable{FF}{<}{<}}
  \TSetTick{cIcro}{1}{\ttable{FT}{<}{\geq}}
  \TSetTick{cIcro}{2}{\ttable{TF}{\geq}{<}}
  \TSetTick{cIcro}{3}{\ttable{TT}{\geq}{\geq}}
  \TSetSortLabel{cIcro}{$\cIcro$}
  \TSort{(4,0.5)}{cIcro}{4}{r}

@@

We first note that in the case where $\cro_3$ is present, we have from \pref{eq:ctx-sigma}:
\begin{align*}
  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \cI_0, \cro_3, \cIcro_{FT} \} \\
  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \cI_1, \cro_3, \cIcro_{TT} \} \\
  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \cI_2, \cro_3, \cIcro_{TT} \}
\end{align*}
We can then deduce the following directions of evolution from \pref{eq:irB}:
\begin{align*}
  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} \\
%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} \\
%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \}
%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
%     \text{(from \pref{eq:B-bounce})}
\end{align*}
          %      eq:inference-edges-ba
Therefore, \pref{eq:edges-inference} gives: $\cI \xrightarrow{1} \N \in \hat{E}_-$.
The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
do not add more information on the influences towards $\N$.
Thus, given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cI}{-}{1}{\N} \in E$.

@@

Let us first focus on the group of regulators $\{ \cro \}$.
Of course, $\forall \cro_i \in \PHl_{\cro}, \ctx_{\{ \cro \}}(\PHetat{\cro_i}) = \{ \cro_i \}$,
and from \pref{eq:irB} and \eqref{eq:fp}:
\begin{align*}
  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_3}) &= \{ \cro_2 \} && \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_2}) &= \{ \cro_1 \} && \text{and} &
    \Phi&(\cro_2) = \emptyset \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_1}) &= \{ \cro_1 \} && \text{and} &
    \Phi&(\cro_1) = \emptyset \\
%     \text{(from \pref{eq:B-no-bounce})} \\
  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_0}) &= \{ \cro_0 \} && \text{and} &
    \PHetat{\cI_2, \cro_0} \in \Phi&(\cro_0)
%     \text{(from \pref{eq:B-no-bounce})}
\end{align*}
We thus cannot infer any local influence from \pref{eq:edges-inference-auto}
by observing the group of regulators $\{ \cro \}$,
because $\Phi(\cro_1) = \Phi(\cro_2) = \emptyset$.
This is due to the fact that all processes of $\cI$ can hit $\cro_1$ and $\cro_2$;
these processes can therefore not be considered stable in any configuration.

Let us now focus instead on the group of regulators $\{ \cI \}$.
One again, we trivially have:
$\forall \cI_i \in \PHl_{\cI}, \forall \cro_j \in \PHl_{\cro},
  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_i, \cro_j}) = \{ \cI_i, \cro_j \}$.
Furthermore, from \pref{eq:irB}, we especially find:
\begin{align*}
%   \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_0}) &= \{ \cro_1 \} &&
%   \text{(from \pref{eq:B-bounce})} \\
%   \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_1}) &= \{ \cro_2 \} &&
%   \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_2}) &= \{ \cro_3 \} \\
%     \text{(from \pref{eq:B-bounce})} \\
  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3}) &= \{ \cro_2 \}
%     \text{(from \pref{eq:B-bounce})}
\end{align*}
Therefore, \pref{eq:edges-inference} gives: $\cro \xrightarrow{3} \cro \in \hat{E}_-$.
% The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
% do not add more information on the influences towards $\N$.
Finally, as no more information is added by observing the other levels of $\cro$ and $\cI$,
and given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cro}{-}{3}{\cro} \in E$.

@@

\begin{table}[t]
\begin{align*}
  K_{\cI, \emptyset} &= \{2\} &
  K_{\N, \emptyset} &= \{1\} \\
  K_{\cI, \{\cro\}} &= \{0\} &
  K_{\N, \{\cro\}} &= \{0\} \\
  K_{\cI, \{\cII\}} &= \{2\} &
  K_{\N, \{\cI\}} &= \{0\} \\
  K_{\cI, \{\cII;cro\}} &= \{2\} &
  K_{\N, \{\cI;cro\}} &= \{0\} \\
  \\
  K_{\cII, \emptyset} &= \{0\} &
  K_{\cro, \emptyset} &= \{3\} \\
  K_{\cII, \{\cro\}} &= \{0\} &
  K_{\cro, \{\cro\}} &= \{2\} \\
  K_{\cII, \{\cI\}} &= \{0\} &
  K_{\cro, \{\cI\}} &= \{0\} \\
  K_{\cII, \{\cI;cro\}} &= \{0\} &
  K_{\cro, \{\cI;cro\}} &= \{2\} \\
  K_{\cII, \{\N\}} &= \{1\} \\
  K_{\cII, \{\N;cro\}} &= \{0\} \\
  K_{\cII, \{\N;cI\}} &= \{0\} \\
  K_{\cII, \{\N;cI;cro\}} &= \{0\}
\end{align*}
\caption{\label{tb:phage-lambda-k}%
  Result of the parameters inference performed on the Process Hitting model
  of the phage lambda immunity response
  taken from \cite{thieffry_dynamical_1995}.
  \todo{Remove the curly brackets $\{\}$?}
}
\end{table}
